---
title: "p8105_hw6_jh4054"
author: "Joy Hsu"
date: "11/24/2018"
output: 
    html_document:
      toc: true
      toc_float: true
      code_folding: show
---

```{r setup, include=FALSE}
library(tidyverse)
library(knitr)

knitr::opts_chunk$set(
  collapse = TRUE)

set.seed(1)

theme_set(theme_bw() + theme(legend.position = "bottom"))
```

### Problem 1

The homicide dataset was collated by Washington Post on 52179 homicide cases in 50 major US cities, from 2007 to 2015. Altogether, 28 states are represented. The raw dimensions are 52179 rows by 12 columns variables. Each case is reported as a distinct row, with demographic information on the victim's name, age, sex, and race. Attributes regarding the homicide incident include geospatial coordinates, reported date, city, and disposition of the police case. 
```{r}
#examine raw dataset
raw_dataset = read_csv("./data/homicide-data.csv")
dim(raw_dataset)

#represented states and cities
raw_dataset %>% 
  distinct(state) %>% 
  nrow()

raw_dataset %>% 
  distinct(city) %>% 
  nrow()
```

```{r}
#load and tidy dataset
homicide_data = read_csv("./data/homicide-data.csv") %>% 
  janitor::clean_names() %>% 
  mutate(city_state = str_c(city, ", ", state)) %>% 
  filter(!(city_state %in% c("Tulsa, AL", "Dallas, TX", "Phoenix, AZ", "Kansas City, MO")))
```

